---
title: "Annual Talking Points QC Documentation"
author: "Amina Chtourou, Gina Sarfaty, Karishma Srikanth"
date: "2/6/2024"
output:
  html_document:
    fig_width: 10
    fig_height: 5.625
    theme: flatly
    dpi: 330
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

  library(tidyverse)
  library(rmarkdown)
  library(glitr)
  library(glamr)
  library(gophr)
  library(systemfonts)
  library(scales)
  library(tidytext)
  library(patchwork)
  library(ggtext)
  library(glue)
  library(janitor)
  library(lubridate)
  library(googlesheets4)
  library(mindthegap)

day <- lubridate::today()
  
```

```{r option, include=FALSE, echo=FALSE}
options(scipen=999)
```

```{r data, include=FALSE, echo=FALSE}
#Current MSD
filepath <- si_path() %>% 
  return_latest("MER_Structured_Datasets_OU_IM_FY22-24")

metadata_msd <- list(curr_pd = "FY23Q4",
                  curr_fy = 2023,
                  curr_fy_lab = "FY23",
                  curr_qtr = 4,
                  source = "FY23Q4c MSD",
                  caption = "Source: FY23Q4c MSD")

# MSD PREP --------------------------------------

#read MSD
df_msd <- read_psd(filepath)

#read NGA genie file
df_nga <- si_path() %>% 
  return_latest("All indicators_FY23Q4_Nigeria") %>% 
  read_csv()

#select only names in MSD
names_to_keep <- names(df_msd)
setdiff(names(df_msd), names(df_nga)) #only qtrs
setdiff(names(df_nga), names(df_msd))

#select all names from NGA file that match MSD (minus qtr1-qtr4)
df_nga_new <- df_nga %>% 
  select(any_of(names_to_keep))

#rbind together removing TZ and NGA for FY23 ONLY
df <- df_msd %>% 
  filter(!(operatingunit=="Nigeria" & fiscal_year=="2023")
         # ,
         # !(operatingunit=="Tanzania" & fiscal_year=="2023")
         ) %>% 
  select(-c(qtr1:qtr4)) %>% 
  rbind(df_nga_new)

# #Q4 TZA Daily Genie 
# df_TZA <- si_path() %>%
#   return_latest("Genie-OUByIMs-Tanzania-Daily-2024-02-07") %>%
#   read_psd()
# 
# #exclude extra columns 
# df_msd_TZA <- df_TZA %>% 
#   select(-c(qtr1:qtr4)) 

#bind TZA data to clean MSD w/ NGA file
df_final <- df 
# %>% 
#   rbind(df_msd_TZA)

# load source functions
#source("Scripts/archive/99_utilities.R")

```

```{r data_prep_functions, include=FALSE, echo=FALSE}

#DREAMS PREP
# Google sheet ID for DREAMS DSNU list
  dsnu_g_id <- "1YaGWvpkiXVPiAwA3KyHwFtbZHrVeCz6DijWPCKpUN-Q"
  
  #MSD path
  dsnu_msd_path <- si_path() %>% 
    return_latest("PSNU_IM_DREAMS_FY22-24")
  
    #import DREAMS DSNU crosswalk
    #recommend adding psnu_uids to this sheet
 dsnu_list <- read_sheet(dsnu_g_id, sheet = "CURRENT (FY23/COP22)")

#rename names
names(dsnu_list)<-c("operatingunit", "psnu", "dsnu", "agencies_FY23")

#import msd and filter to current year
df_msd_dreams <- read_psd(dsnu_msd_path) %>% 
  filter(fiscal_year == metadata_msd$curr_fy)

msd_dsnu_xwalk <- df_msd_dreams %>% 
  count(operatingunit, operatingunituid, psnu, psnuuid, cop22_psnu, cop22_psnuuid, dsnu, dsnuuid) %>% 
  mutate(raised_lvl = ifelse(psnuuid != cop22_psnuuid, TRUE, FALSE)) %>% 
  mutate(dsnu_new = case_when(operatingunit != "Uganda" & raised_lvl == TRUE ~ cop22_psnu,
                          TRUE ~ dsnu),
         dsnuuid_new = case_when(operatingunit != "Uganda" & raised_lvl == TRUE ~ cop22_psnuuid,
                          TRUE ~ dsnuuid))

#join the new xwalk back to msd and then join the internal dsnu list to the msd
df_dreams_all <- df_msd_dreams %>% 
  left_join(msd_dsnu_xwalk) %>% 
  left_join(dsnu_list, by=c("operatingunit", "cop22_psnu" = "psnu", "dsnu_new"= "dsnu"))

#LOCAL PARTNER PREP
#Read in the google sheet hyperfile with local partner
sheet_id <- "1MQviknJkJDttGdNEJeNaYPKmHCw6BqPuJ0C5cslV5IE"

df_partner <- read_sheet(sheet_id, sheet = "MechID-PartnerType", range = "A:B") %>% 
  clean_names() %>% 
  rename(mech_code = mechanism_id) %>% 
  mutate(mech_code = as.character(mech_code),
         partner_type = case_when(partner_type == "Regional" ~ "Local",
                                  partner_type == "TBD Local" ~ "Local", TRUE ~ partner_type))

#FUNCTIONS----------------------------------------------------------------------

#clean number function
clean_number <- function(x, digits = 0){
  dplyr::case_when(x >= 1e9 ~ glue("{round(x/1e9, digits)} billion"),
                   x >= 1e6 ~ glue("{round(x/1e6, digits)} million"),
                   x >= 1e3 ~ glue("{round(x/1e3, digits)} thousand"),
                   TRUE ~ glue("{x}"))
}

#LP Function
#result_type = value or share
grab_lp_results_clean <- function(indic, result_type) {
  
  df_munge <- suppressMessages(df_final %>% 
                                 # bind_rows(df_arch) %>%
                                 left_join(df_partner, by = c("mech_code")) %>%
                                 filter(funding_agency == "USAID",
                                        indicator == indic,
                                        standardizeddisaggregate == "Total Numerator") %>% 
                                 group_by(funding_agency, fiscal_year, indicator, partner_type) %>% 
                                 summarise(across(cumulative, sum, na.rm = TRUE)) %>%
                                 ungroup() %>% 
                                 filter(partner_type != "TBD") %>%
                                 #filter(fiscal_year != 2022) %>%
                                 pivot_wider(names_from = partner_type, values_from = cumulative) %>%
                                 group_by(fiscal_year) %>%
                                 mutate(Total = International + Local,
                                        share = Local / Total)  %>%
                                 pivot_longer(cols = International:Total, names_to = "partner_type"))
  
  title_info_lp <- df_munge %>% 
    filter(partner_type == "Local", fiscal_year == metadata_msd$curr_fy) %>% 
    select(fiscal_year, indicator, value, share) %>% 
    mutate(
      value = value %>% clean_number(1),#change to 1 if you want 1 decimal accuracy
      share = percent(round(share, 2))) %>% 
    pull(result_type)
  
  return(title_info_lp)
}

grab_lp_results <- function(indic, result_type) {
  
  df_munge <- suppressMessages(df_final %>% 
                                 # bind_rows(df_arch) %>%
                                 left_join(df_partner, by = c("mech_code")) %>%
                                 filter(funding_agency == "USAID",
                                        indicator == indic,
                                        standardizeddisaggregate == "Total Numerator") %>% 
                                 group_by(funding_agency, fiscal_year, indicator, partner_type) %>% 
                                 summarise(across(cumulative, sum, na.rm = TRUE)) %>%
                                 ungroup() %>% 
                                 filter(partner_type != "TBD") %>%
                                 #filter(fiscal_year != 2022) %>%
                                 pivot_wider(names_from = partner_type, values_from = cumulative) %>%
                                 group_by(fiscal_year) %>%
                                 mutate(Total = International + Local,
                                        share = Local / Total)  %>%
                                 pivot_longer(cols = International:Total, names_to = "partner_type"))
  
  title_info_lp <- df_munge %>% 
    filter(partner_type == "Local", fiscal_year == metadata_msd$curr_fy) %>% 
    select(fiscal_year, indicator, value, share) %>% 
    mutate(
      value = value,#change to 1 if you want 1 decimal accuracy
      share = percent(round(share, 2))) %>% 
    pull(result_type) %>%
  
  return(title_info_lp)
}


#key pops function clean
pull_kp_clean <- function(indic) {
  kp_val <- df_final %>%
    filter(funding_agency == "USAID",
           str_detect(standardizeddisaggregate, "KeyPop(?!\\/Status)"),
           indicator %in% c("PrEP_NEW", "TX_CURR", "KP_PREV"),
           fiscal_year == metadata_msd$curr_fy) %>%
    count(indicator, wt = cumulative) %>%
    mutate(n = clean_number(n, 1)) %>% 
    pivot_wider(names_from = indicator, values_from = "n") %>% 
    pull(indic)
  
  return(kp_val)
}

#key pops function complete
pull_kp <- function(indic) {
  kp_val <- df_final %>%
    filter(funding_agency == "USAID",
           str_detect(standardizeddisaggregate, "KeyPop(?!\\/Status)"),
           indicator %in% c("PrEP_NEW", "TX_CURR", "KP_PREV"),
           fiscal_year == metadata_msd$curr_fy) %>%
    mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                             operatingunit == "Tanzania" ~ "Tanzania",
                             TRUE ~ "All other OUs")) %>%
    count(fiscal_year, indicator, ou_qc, wt = cumulative) %>%
    pivot_wider(names_from = indicator, values_from = "n") %>%
    select(fiscal_year, ou_qc, all_of(indic))

  return(kp_val)
}


load_secrets()

```

Last Updated: `r day`

## **BLUF**

## **Epidemic Control Checks Documentation**

The epi control section has been verified and confirmed in [this document](https://docs.google.com/document/d/11ezAbTw6eWkBXu-NZrINt3iSNcPYhoEMswJoZtPOu9w/edit).

## **Prevention Talking Points**

### TP #5: PrEP and local partner share

```{r prevention_prep, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
library(gt)
library(gtExtras)

prep_qc <- df_final %>%
  filter(funding_agency == "USAID",
         standardizeddisaggregate %in% c("Total Numerator"),
         indicator %in% c("PrEP_NEW"),
         fiscal_year == metadata_msd$curr_fy) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                        #   operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>% 
  group_by(fiscal_year, indicator, funding_agency, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() %>% 
  mutate(fiscal_year = as.character(fiscal_year))

prep_qc %>% 
  #filter(`OU category` != "Total") %>% 
  gt::gt() %>% 
  fmt_number(columns = c(5), 
             decimals = 0) %>% 
  grand_summary_rows(
    columns = where(is.numeric),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #5: PrEP_NEW Initiations" %>% toupper())) %>% 
  gt_theme_nytimes() 

```

```{r prep_lp_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

df_prep_lp_qc <- df_final %>%
    left_join(df_partner, by = c("mech_code")) %>%
    filter(funding_agency == "USAID",
           indicator == "PrEP_NEW",
           standardizeddisaggregate == "Total Numerator") %>%
    mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                          # operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(funding_agency, fiscal_year, indicator, partner_type, ou_qc) %>%
  summarise(across(cumulative, sum, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(partner_type != "TBD") %>%
  pivot_wider(names_from = partner_type, values_from = cumulative) %>%
  group_by(fiscal_year) %>%
  mutate(Total = International + Local,
         share = Local / Total)  %>%
  pivot_longer(cols = International:Total, names_to = "partner_type") %>% 
  ungroup()

qc_prep_lp <- df_prep_lp_qc %>% 
  filter(partner_type %in% c("Local", "Total"), fiscal_year == metadata_msd$curr_fy) %>% 
  pivot_wider(names_from = "partner_type", values_from = "value")
```

```{r prep_lp_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

qc_prep_lp %>% 
  mutate(fiscal_year = as.character(fiscal_year)) %>% 
  relocate(share, .after = Total) %>% 
  gt() %>% 
  fmt_number(columns = c(5,6), 
             decimals = 0) %>% 
  fmt_percent(columns = 7, 
              decimals = 1) %>% 
  grand_summary_rows(
    columns = c(5,6),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #5: Local Partner PrEP share" %>% toupper())) %>% 
  gt_theme_nytimes() 

```

### TP #7: VMMC_CIRC

```{r vmmc_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
vmmc_qc <- df_final %>% 
  filter(indicator == "VMMC_CIRC",
         standardizeddisaggregate == "Total Numerator",
         funding_agency == "USAID",
         fiscal_year == metadata_msd$curr_fy) %>% 
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           #operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(fiscal_year, funding_agency, ou_qc) %>% 
  summarise_at(vars(cumulative, targets), sum, na.rm = TRUE) %>%
  ungroup() %>% 
  mutate(achv = cumulative/targets)

```

```{r vmmc_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

vmmc_qc %>% 
  gt() %>% 
  fmt_number(columns = c(4,5), 
             decimals = 0) %>% 
  fmt_percent(columns = 6, 
              decimals = 1) %>% 
  grand_summary_rows(
    columns = c(4,5),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #7: FY23 VMMC_CIRC" %>% toupper())) %>% 
  gt_theme_nytimes() 

```

------------------------------------------------------------------------

## **Addressing the 95s: Testing, Treatment, and Viral Suppression - Talking Points**

### TP #8: FY23 USAID HTS_TST and HTS_POS

```{r hts_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
df_hts_qc <- df_final %>% 
  filter(indicator %in% c("HTS_TST", "HTS_TST_POS"),
         fiscal_year %in% c(metadata_msd$curr_fy -1, metadata_msd$curr_fy),
         standardizeddisaggregate == "Total Numerator") %>% 
 #   bind_rows(df_hts %>% mutate(funding_agency = "PEPFAR")) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           #operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
    group_by(fiscal_year, funding_agency, indicator, ou_qc) %>%
    summarise(across(c(cumulative), sum, na.rm = TRUE)) %>%
    ungroup() %>%
    filter(funding_agency %in% c("USAID")) %>%
    pivot_wider(names_from = indicator, values_from = cumulative) 
```

```{r hts_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
df_hts_qc %>% 
  gt(groupname_col = "fiscal_year") %>% 
  fmt_number(columns = c(4,5), 
             decimals = 0) %>% 
  # fmt_percent(columns = 6, 
  #             decimals = 1) %>% 
  summary_rows(
    columns = c(4,5),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #8: FY23 HTS and HTS_POS" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

### TP #9: HTS_SELF

```{r hts_self_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
df_hts_self_qc <- df_final %>% 
  filter(indicator %in% c("HTS_SELF"),
         fiscal_year %in% c(metadata_msd$curr_fy -1, metadata_msd$curr_fy),
         standardizeddisaggregate == "Total Numerator") %>% 
 #   bind_rows(df_hts %>% mutate(funding_agency = "PEPFAR")) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                         #  operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
    group_by(fiscal_year, funding_agency, indicator, ou_qc) %>%
    summarise(across(c(cumulative), sum, na.rm = TRUE)) %>%
    ungroup() %>%
    filter(funding_agency %in% c("USAID")) %>%
    pivot_wider(names_from = indicator, values_from = cumulative) 
```

```{r hts_self_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
df_hts_self_qc %>% 
  gt(groupname_col = "fiscal_year") %>% 
  fmt_number(columns = c(4), 
             decimals = 0) %>% 
  # fmt_percent(columns = 6, 
  #             decimals = 1) %>% 
  summary_rows(
    columns = c(4),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #9: FY23 HTS_SELF" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

### TP #10:  FY23 USAID Youth (10-24) HTS and HTS_POS

```{r hts_youth_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
#youth testing 
df_hts_youth <- df_final %>%
  filter(target_age_2024 %in% c("10-14", "15-24"),
         indicator %in% c("HTS_TST", "HTS_TST_POS"),
         standardizeddisaggregate == "Modality/Age/Sex/Result",
         funding_agency == "USAID",
         fiscal_year == metadata_msd$curr_fy) %>% 
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                         #  operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs"),
         age_band = "10-24") %>%
  group_by(fiscal_year, funding_agency, age_band, indicator, ou_qc) %>%
  summarise_at(vars(cumulative, targets), sum, na.rm = TRUE) %>%
  ungroup() %>% 
  mutate(achv = cumulative/targets) 
```

```{r hts_youth_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
df_hts_youth %>% 
  gt() %>% 
  tab_row_group(
    label = "HTS_TST_POS",
    rows = indicator %in% c("HTS_TST_POS")
  ) %>%
  tab_row_group(
    label = "HTS_TST",
    rows = indicator %in% c("HTS_TST")
  ) %>% 
  fmt_number(columns = c(6,7), 
             decimals = 0) %>% 
  fmt_percent(columns = 8,
              decimals = 1) %>%
 summary_rows(
    columns = c(6,7),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #19: FY23 YOUTH HTS and HTS_POS" %>% toupper())) %>% 
  gt_theme_nytimes() 
```


------------------------------------------------------------------------

### TP #7: OVC

```{r ovc_serv_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

ovc_serv_qc <- df_final %>%
  filter(indicator %in% c("OVC_SERV"),
         standardizeddisaggregate %in% c("Total Numerator"),
         fiscal_year == metadata_msd$curr_fy,
         funding_agency == "USAID")%>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>% 
  group_by(fiscal_year, indicator, funding_agency, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() %>% 
  mutate(fiscal_year = as.character(fiscal_year))
```

```{r ovc_serv_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
ovc_serv_qc %>% 
  #filter(`OU category` != "Total") %>% 
  gt::gt() %>% 
  fmt_number(columns = c(5), 
             decimals = 0) %>% 
  grand_summary_rows(
    columns = where(is.numeric),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #7: OVC_SERV" %>% toupper())) %>% 
  gt_theme_nytimes() 

```


```{r ovc_hivstat_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

ovc_hivstat_qc <- df_final %>%
  filter(indicator %in% c("OVC_HIVSTAT_POS"),
        standardizeddisaggregate %in% c("Total Numerator"),
         fiscal_year == metadata_msd$curr_fy,
         funding_agency == "USAID")%>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>% 
  group_by(fiscal_year, indicator, funding_agency, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() %>% 
  mutate(fiscal_year = as.character(fiscal_year))
```

```{r ovc_hivstat_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
ovc_hivstat_qc %>% 
  #filter(`OU category` != "Total") %>% 
  gt::gt() %>% 
  fmt_number(columns = c(5), 
             decimals = 0) %>% 
  grand_summary_rows(
    columns = where(is.numeric),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #7: OVC_HIVSTAT_POS" %>% toupper())) %>% 
  gt_theme_nytimes() 

```


### TP #8: OVC+DREAMS PREVENTION
```{r ovc_preventative_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

ovc_prev_qc <- df_final %>%
  filter(indicator %in% c("OVC_SERV"),
        standardizeddisaggregate %in% c("Age/Sex/Preventive"),
         fiscal_year == metadata_msd$curr_fy,
         funding_agency == "USAID",
        ageasentered == "10-14")%>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>% 
  group_by(fiscal_year, indicator,ageasentered,funding_agency, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() %>% 
  mutate(fiscal_year = as.character(fiscal_year))
```

```{r ovc_preventative_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
ovc_prev_qc %>% 
  #filter(`OU category` != "Total") %>% 
  gt::gt() %>% 
  fmt_number(columns = c(5), 
             decimals = 0) %>% 
  grand_summary_rows(
    columns = where(is.numeric),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #7: OVC_SERV Prev 10-14" %>% toupper())) %>% 
  gt_theme_nytimes() 

```

## **Prevention Talking Points**

### TP #9: PrEP

```{r prevention_prep_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
library(gt)
library(gtExtras)

prep_qc <- df_final %>%
  filter(funding_agency == "USAID",
         standardizeddisaggregate %in% c("Total Numerator"),
         indicator %in% c("PrEP_NEW"),
         fiscal_year == metadata_msd$curr_fy) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>% 
  group_by(fiscal_year, indicator, funding_agency, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() %>% 
  mutate(fiscal_year = as.character(fiscal_year))
```

```{r prevention_prep_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

prep_qc %>% 
  #filter(`OU category` != "Total") %>% 
  gt::gt() %>% 
  fmt_number(columns = c(5), 
             decimals = 0) %>% 
  grand_summary_rows(
    columns = where(is.numeric),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #9: PrEP_NEW Initiations" %>% toupper())) %>% 
  gt_theme_nytimes() 

```

### TP #10: PrEP Local Partner Share

```{r prep_lp_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

df_prep_lp_qc <- df_final %>%
    left_join(df_partner, by = c("mech_code")) %>%
    filter(funding_agency == "USAID",
           indicator == "PrEP_NEW",
           standardizeddisaggregate == "Total Numerator") %>%
    mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(funding_agency, fiscal_year, indicator, partner_type, ou_qc) %>%
  summarise(across(cumulative, sum, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(partner_type != "TBD") %>%
  pivot_wider(names_from = partner_type, values_from = cumulative) %>%
  group_by(fiscal_year) %>%
  mutate(Total = International + Local,
         share = Local / Total)  %>%
  pivot_longer(cols = International:Total, names_to = "partner_type") %>% 
  ungroup()

qc_prep_lp <- df_prep_lp_qc %>% 
  filter(partner_type %in% c("Local", "Total"), fiscal_year == metadata_msd$curr_fy) %>% 
  pivot_wider(names_from = "partner_type", values_from = "value")
```

```{r prep_lp_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

qc_prep_lp %>% 
  mutate(fiscal_year = as.character(fiscal_year)) %>% 
  relocate(share, .after = Total) %>% 
  gt() %>% 
  fmt_number(columns = c(5,6), 
             decimals = 0) %>% 
  fmt_percent(columns = 7, 
              decimals = 1) %>% 
  grand_summary_rows(
    columns = c(5,6),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #10: Local Partner PrEP share" %>% toupper())) %>% 
  gt_theme_nytimes() 

```

### TP #12: KP PrEP Initiations

```{r prep_kp_agyw_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

#AGYW prep
prep_agyw_val <- df_final %>%
  filter(funding_agency == "USAID",
         standardizeddisaggregate %in% c("Age/Sex"),
         indicator %in% c("PrEP_NEW"),
         fiscal_year == metadata_msd$curr_fy,
         sex == "Female",
         age_2019 %in% c("10-14","15-19", "20-24")
  ) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>% 
  group_by(fiscal_year, funding_agency, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  rename(`PrEP_NEW AGYW` = cumulative)

prep_kp_val <- df_final %>%
    filter(funding_agency == "USAID",
           str_detect(standardizeddisaggregate, "KeyPop(?!\\/Status)"),
           indicator %in% c("PrEP_NEW"),
           fiscal_year == metadata_msd$curr_fy) %>%
    mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                             operatingunit == "Tanzania" ~ "Tanzania",
                             TRUE ~ "All other OUs")) %>%
    count(fiscal_year, indicator, ou_qc, wt = cumulative) %>%
    pivot_wider(names_from = indicator, values_from = "n") %>%
    select(fiscal_year, ou_qc, `PrEP_NEW`)
```


```{r prep_kp_agyw_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

prep_kp_val %>% 
  rename(`PrEP_NEW KP` = PrEP_NEW ) %>% 
  left_join(prep_agyw_val, by = c("fiscal_year", "ou_qc")) %>% 
  select(fiscal_year, funding_agency, ou_qc, `PrEP_NEW KP`, `PrEP_NEW AGYW`) %>% 
  gt() %>% 
  fmt_number(columns = c(4,5), 
             decimals = 0) %>% 
  grand_summary_rows(
    columns = c(4,5),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #12: KP and AGYW PrEP Initiations" %>% toupper())) %>% 
  gt_theme_nytimes() 

```

### TP #13: VMMC_CIRC

```{r vmmc_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
vmmc_qc <- df_final %>% 
  filter(indicator == "VMMC_CIRC",
         standardizeddisaggregate == "Total Numerator",
         funding_agency == "USAID",
         fiscal_year == metadata_msd$curr_fy) %>% 
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(fiscal_year, funding_agency, ou_qc) %>% 
  summarise_at(vars(cumulative, targets), sum, na.rm = TRUE) %>%
  ungroup() %>% 
  mutate(achv = cumulative/targets)

```

```{r vmmc_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

vmmc_qc %>% 
  gt() %>% 
  fmt_number(columns = c(4,5), 
             decimals = 0) %>% 
  fmt_percent(columns = 6, 
              decimals = 1) %>% 
  grand_summary_rows(
    columns = c(4,5),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #13: FY23 VMMC_CIRC" %>% toupper())) %>% 
  gt_theme_nytimes() 

```

### TP #14: USAID AGYW_PREV Total DREAMS 

```{r total_agyw_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
#For total DREAMS programming, use all 4 disaggs + ages 10-29
viz_package <- df_dreams_all %>%
  filter(indicator=="AGYW_PREV",
         standardizeddisaggregate %in% c("Age/Sex/Time/Complete+",
                                         "Age/Sex/Time/Complete",
                                         "Age/Sex/Time/Started",
                                         "Age/Sex/Time/Incomplete"),
         numeratordenom=="D",
         age_2019 %in% c("10-14","15-19","20-24","25-29"),
         sex=="Female")


#create shares by USAID and non-USAID contributing DSNUs
dreams_overall_val <- viz_package %>% 
  mutate(usaid_dsnu = ifelse(str_detect(agencies_FY23, "USAID"), "USAID", "non-USAID")) %>% 
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(indicator, fiscal_year, usaid_dsnu, ou_qc) %>% 
  summarize(across(cumulative, \(x) sum(x, na.rm = TRUE)),.groups="drop") %>% 
  pivot_wider(names_from = "usaid_dsnu", values_from = 'cumulative') %>% 
  mutate(total = USAID + `non-USAID`,
         usaid_share = USAID/total)
```

```{r total_agyw_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
dreams_overall_val %>% 
  select(-c(`non-USAID`)) %>% 
  gt() %>% 
  fmt_number(columns = c(4,5), 
             decimals = 0) %>% 
  fmt_percent(columns = 6, 
              decimals = 1) %>% 
  grand_summary_rows(
    columns = c(4,5),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #14: USAID AGYW_PREV Total DREAMS programming" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

### TP #16: USAID AGYW_PREV by Economic/Education Disaggregates

```{r agyw_edu_econ_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
#service type disagg
viz_service<- df_dreams_all %>% filter(indicator=="AGYW_PREV", 
                                       standardizeddisaggregate %in% c("ComprehensiveEconomicStrengthening",
                                                                       "EducationSupport"),
                                       numeratordenom=="D")

## Filter for education and econ disaggs + USAID
df_serv_type <- viz_service %>%
  filter(str_detect(agencies_FY23, "USAID")) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(fiscal_year, standardizeddisaggregate, ou_qc) %>%
  summarize(across(cumulative, \(x) sum(x, na.rm = TRUE)),.groups="drop") %>%
  pivot_wider(names_from = "standardizeddisaggregate", values_from = "cumulative") 
```

```{r agyw_edu_econ_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
df_serv_type %>% 
  gt() %>% 
  fmt_number(columns = c(3,4), 
             decimals = 0) %>% 
  # fmt_percent(columns = 6, 
  #             decimals = 1) %>% 
  grand_summary_rows(
    columns = c(3,4),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #16: USAID AGYW_PREV by Economic/Education Disaggs" %>% toupper())) %>% 
  gt_theme_nytimes() 
```


## **Test and Diagnosis Talking Points**

### TP #17: FY23 USAID HTS and HTS_POS

```{r hts_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
df_hts_qc <- df_final %>% 
  filter(indicator %in% c("HTS_TST", "HTS_TST_POS"),
         fiscal_year == metadata_msd$curr_fy,
         standardizeddisaggregate == "Total Numerator") %>% 
 #   bind_rows(df_hts %>% mutate(funding_agency = "PEPFAR")) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
    group_by(fiscal_year, funding_agency, indicator, ou_qc) %>%
    summarise(across(c(cumulative), sum, na.rm = TRUE)) %>%
    ungroup() %>%
    filter(funding_agency %in% c("USAID")) %>%
    pivot_wider(names_from = indicator, values_from = cumulative) 
```

```{r hts_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
df_hts_qc %>% 
  gt() %>% 
  fmt_number(columns = c(4,5), 
             decimals = 0) %>% 
  # fmt_percent(columns = 6, 
  #             decimals = 1) %>% 
  grand_summary_rows(
    columns = c(4,5),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #17: FY23 HTS and HTS_POS" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

### TP #18: HTS_SELF

```{r hts_self_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
df_hts_self_qc <- df_final %>% 
  filter(indicator %in% c("HTS_SELF"),
         fiscal_year == metadata_msd$curr_fy,
         standardizeddisaggregate == "Total Numerator") %>% 
 #   bind_rows(df_hts %>% mutate(funding_agency = "PEPFAR")) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
    group_by(fiscal_year, funding_agency, indicator, ou_qc) %>%
    summarise(across(c(cumulative), sum, na.rm = TRUE)) %>%
    ungroup() %>%
    filter(funding_agency %in% c("USAID")) %>%
    pivot_wider(names_from = indicator, values_from = cumulative) 
```

```{r hts_self_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
df_hts_self_qc %>% 
  gt() %>% 
  fmt_number(columns = c(4), 
             decimals = 0) %>% 
  # fmt_percent(columns = 6, 
  #             decimals = 1) %>% 
  grand_summary_rows(
    columns = c(4),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #18: FY23 HTS_SELF" %>% toupper())) %>% 
  gt_theme_nytimes() 
```


### TP #21:  Local Partner HTS share

```{r hts_lp_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
hts_lp_qc <- df_final %>%
  left_join(df_partner, by = c("mech_code")) %>%
  filter(funding_agency == "USAID",
         indicator == "HTS_TST",
         standardizeddisaggregate == "Total Numerator") %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(funding_agency, fiscal_year, indicator, partner_type, ou_qc) %>%
  summarise(across(cumulative, sum, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(partner_type != "TBD") %>%
  pivot_wider(names_from = partner_type, values_from = cumulative) %>%
  group_by(fiscal_year) %>%
  mutate(Total = International + Local,
         share = Local / Total)  %>%
  pivot_longer(cols = International:Total, names_to = "partner_type") %>% 
  ungroup()

qc_hts_lp <- hts_lp_qc %>% 
  filter(partner_type %in% c("Local", "Total"), fiscal_year == metadata_msd$curr_fy) %>% 
  pivot_wider(names_from = "partner_type", values_from = "value")
```

```{r hts_lp_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
qc_hts_lp %>% 
  mutate(fiscal_year = as.character(fiscal_year)) %>% 
  relocate(share, .after = Total) %>% 
  gt() %>% 
  fmt_number(columns = c(5,6), 
             decimals = 0) %>% 
  fmt_percent(columns = 7, 
              decimals = 1) %>% 
  grand_summary_rows(
    columns = c(5,6),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #21: Local Partner HTS share" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

## **Treatment and Viral Suppression Talking Points**


### TP #24:  VERTICAL TRANSMISSION

```{r pmtct_stat_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
#pmtct_stat
pmtct_stat_qc<-df_final %>% 
  filter(indicator=="PMTCT_STAT",
         standardizeddisaggregate %in% c("Total Numerator","Total Denominator"),
         funding_agency=="USAID",
         fiscal_year == metadata_msd$curr_fy) %>% 
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>% 
  group_by(fiscal_year, funding_agency,standardizeddisaggregate, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() %>% 
  spread(standardizeddisaggregate,cumulative) %>% 
  mutate(prct=(`Total Numerator`/`Total Denominator`))
  
```

```{r pmtct_stat_table, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
pmtct_stat_qc %>% 
  gt() %>% 
  tab_row_group(
    label = "PMTCT_STAT",
    rows = funding_agency %in% c("USAID")
  ) %>%
  fmt_number(columns = c(6,6), 
             decimals = 0) %>% 
  fmt_percent(columns = 6,
              decimals = 1) %>%
 summary_rows(
    columns = c(4,5),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #24: FY23 PMTCT STAT N/D" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

```{r pmtct_post_anc1_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
#pmtct_post_anc_test

pmtct_anc_test_change_qc<-df_final %>% 
   filter(indicator=="HTS_TST",
         standardizeddisaggregate=="Modality/Age/Sex/Result",
         funding_agency=="USAID",
         fiscal_year %in% c("2022","2023"),
         modality=="Post ANC1") %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>% 
  group_by(fiscal_year, funding_agency,modality,ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() %>% 
  spread(fiscal_year,cumulative) %>% 
  mutate(prct_change = (`2022` / `2023`))

  
```

```{r pmtct_post_anc1_table, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
pmtct_anc_test_change_qc %>% 
  gt() %>% 
  tab_row_group(
    label = "POST ANC1 TESTING",
    rows = funding_agency %in% c("USAID")
  ) %>%
  fmt_number(columns = c(6,6), 
             decimals = 0) %>% 
  fmt_percent(columns = 6,
              decimals = 1) %>%
 summary_rows(
    columns = c(4,5),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #24: FY22 & 23 POST ANC1 TESTING" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

```{r pmtct_art_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
#pmtct_art

pmtct_art_qc<-df_final %>% 
   filter(indicator=="PMTCT_ART",
         standardizeddisaggregate %in% c("Total Numerator","Total Denominator"),
         funding_agency=="USAID",
         fiscal_year == metadata_msd$curr_fy) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>% 
  group_by(fiscal_year, funding_agency,numeratordenom,ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() %>% 
  spread(numeratordenom,cumulative) %>% 
  mutate(Percent = N/D)

  
```

```{r pmtct_art_table, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
pmtct_art_qc %>% 
  gt() %>% 
  tab_row_group(
    label = "PMTCT_ART",
    rows = funding_agency %in% c("USAID")
  ) %>%
  fmt_number(columns = c(6,6), 
             decimals = 0) %>% 
  fmt_percent(columns = 6,
              decimals = 1) %>%
 summary_rows(
    columns = c(4,5),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #24: FY23 PMTCT_ART" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

```{r pmtct_vls_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
#pmtct_VLS

pmtct_vls_qc<-df_final %>% 
   filter(indicator=="TX_PVLS",
         standardizeddisaggregate =="PregnantOrBreastfeeding/Indication/HIVStatus",
         grepl("Pregnant", otherdisaggregate),
         funding_agency=="USAID",
         fiscal_year == metadata_msd$curr_fy) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>% 
  group_by(fiscal_year,funding_agency,numeratordenom,ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() %>% 
  spread(numeratordenom,cumulative) %>% 
  mutate(Percent = N/D)

  
```

```{r pmtct_vls_table, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
pmtct_vls_qc %>% 
  gt() %>% 
  tab_row_group(
    label = "PMTCT VLS",
    rows = funding_agency %in% c("USAID")
  ) %>%
  fmt_number(columns = c(6,6), 
             decimals = 0) %>% 
  fmt_percent(columns = 6,
              decimals = 1) %>%
 summary_rows(
    columns = c(4,5),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #24: FY23 PMTCT VLS Among Preg. Women" %>% toupper())) %>% 
  gt_theme_nytimes() 

```
### TP #29:  KP_PREV

```{r kp_prev_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
kp_prev_val <- df_final %>%
  filter(funding_agency == "USAID",
         str_detect(standardizeddisaggregate, "KeyPop(?!\\/Status)"),
         indicator %in% c("KP_PREV"),
         fiscal_year == metadata_msd$curr_fy) %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  count(fiscal_year, funding_agency, indicator, ou_qc, wt = cumulative) %>%
  pivot_wider(names_from = indicator, values_from = "n") %>%
  select(fiscal_year, funding_agency, ou_qc, `KP_PREV`)
```

```{r kp_prev_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}

kp_prev_val %>% 
  gt() %>% 
  fmt_number(columns = c(4), 
             decimals = 0) %>% 
  # fmt_percent(columns = 7, 
  #             decimals = 1) %>% 
  grand_summary_rows(
    columns = c(4),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #29: KP_PREV" %>% toupper())) %>% 
  gt_theme_nytimes() 
```


### TP #31:  CX_CA

```{r prep_cxca_df, echo=FALSE, include = FALSE, warning=FALSE, message=FALSE}

df_arch1 <- si_path() %>% 
  return_latest("OU_IM_FY15") %>% 
  read_psd()

df_arch <- df_arch1 %>% 
  select(-c(qtr1:qtr4)) 

# MUNGE -----------------------------------------------------------------

#bind archived + current MSD and filter 
df_bind <- df_final %>%
  bind_rows(df_arch) %>% 
  dplyr::filter(!(fiscal_year %in% c("2024", "2015", "2016", "2017")))%>% 
  dplyr::filter(funding_agency %in% "USAID")
```

```{r cxca_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
cxca_scrn_gofurther_val <- df_bind %>% 
  filter(indicator %in% c("CXCA_SCRN"),
         standardizeddisaggregate %in% c("Total Numerator"),
         #  trendscoarse == "<18",
         fiscal_year %in% c("2018", "2019", "2020", "2021", "2022", "2023"),
         funding_agency == "USAID",
         country %in% c("Zimbabwe", "Botswana", "Eswatini", "Lesotho",
                        "Malawi", "Mozambique", "Namibia", "Zambia")) %>%
  group_by(funding_agency) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  mutate(ou_qc = "Go Futher Countries") %>% 
  select(funding_agency, ou_qc, cumulative)

cxca_scrn_val <- df_bind %>% 
  filter(indicator %in% c("CXCA_SCRN"),
         standardizeddisaggregate %in% c("Total Numerator"),
         #  trendscoarse == "<18",
         fiscal_year %in% c("2018", "2019", "2020", "2021", "2022", "2023"),
         funding_agency == "USAID") %>% 
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(funding_agency, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") 
```

```{r cxca_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
cxca_scrn_gofurther_val %>% 
  rbind(cxca_scrn_val) %>% 
  mutate(ou_qc = case_when(ou_qc == "All other OUs" ~ "All other OUs minus NGA/TZA",
                           ou_qc == "Go Futher Countries" ~ "Only Go Futher Countries",
                           TRUE ~ ou_qc)) %>% 
  gt() %>% 
  tab_row_group(
    label = "Go Further Countries",
    rows = ou_qc %in% c("Only Go Futher Countries")
  ) %>%
  tab_row_group(
    label = " ",
    rows = ou_qc %ni% c("Only Go Futher Countries")
  ) %>% 
  fmt_number(columns = c(3), 
             decimals = 0) %>% 
  summary_rows(
    columns = c(3),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #31: CXCA_SCRN Cumulative (2018-2023)" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

```{r cxca_pos_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
cxca_scrn_pos_val <- df_bind %>% 
  filter(indicator %in% c("CXCA_SCRN_POS", "CXCA_TX"),
         standardizeddisaggregate %in% c("Total Numerator"),
         #  trendscoarse == "<18",
         fiscal_year %in% c("2018", "2019", "2020", "2021", "2022", "2023"),
         funding_agency == "USAID") %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(indicator, funding_agency, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  pivot_wider(names_from = "indicator", values_from = "cumulative") %>% 
  mutate(tx_prct = CXCA_TX/CXCA_SCRN_POS)
```

```{r cxca_pos_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
cxca_scrn_pos_val %>% 
  gt() %>% 
  fmt_number(columns = c(3,4), 
             decimals = 0) %>% 
  fmt_percent(columns = 5,
              decimals = 1) %>%
  grand_summary_rows(
    columns = c(3,4),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #31: CXCA_SCRN_POS and CXCA_TX Cumulative (FY18-23)" %>% toupper())) %>% 
  gt_theme_nytimes() 
```
### TP #32:  TB

```{r tb_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
tb_stat_qc <- df_final %>% 
  clean_indicator() %>% 
  filter(indicator %in% c("TB_STAT", "TB_STAT_D", "TB_STAT_POS", "TB_ART") ,
         standardizeddisaggregate %in% c("Total Numerator", "Total Denominator"),
         funding_agency == "USAID",
         fiscal_year == metadata_msd$curr_fy) %>% 
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(fiscal_year, funding_agency, indicator, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() %>% 
  pivot_wider(names_from = "indicator", values_from = "cumulative") %>% 
  mutate(`TB_STAT Coverage` = TB_STAT/TB_STAT_D,
         `TB_ART Coverage` = TB_STAT_POS/TB_ART)
```

```{r tb_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
tb_stat_qc %>% 
  relocate(TB_ART, .after = TB_STAT_POS) %>% 
    gt() %>% 
  fmt_number(columns = c(4:7), 
             decimals = 0) %>% 
  fmt_percent(columns = c(8,9),
              decimals = 1) %>%
  grand_summary_rows(
    columns = c(4:7),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #32: TB_STAT, TB_STAT_POS, and TB_ART" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

### TP #33:  GENDER

```{r gend_qc, echo=TRUE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
#total value from MSD
gbv_val_clean <- df_final %>% 
  filter(indicator == "GEND_GBV",
         standardizeddisaggregate == "Total Numerator",
         funding_agency == "USAID",
         fiscal_year == metadata_msd$curr_fy) %>% 
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(fiscal_year, funding_agency, ou_qc) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
  ungroup() 

#Local partner share
qc_gend_lp <- df_final %>%
  left_join(df_partner, by = c("mech_code")) %>%
  filter(funding_agency == "USAID",
         indicator == "GEND_GBV",
         standardizeddisaggregate == "Total Numerator") %>%
  mutate(ou_qc = case_when(operatingunit == "Nigeria" ~ "Nigeria",
                           operatingunit == "Tanzania" ~ "Tanzania",
                           TRUE ~ "All other OUs")) %>%
  group_by(funding_agency, fiscal_year, indicator, partner_type, ou_qc) %>%
  summarise(across(cumulative, sum, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(partner_type != "TBD") %>%
  pivot_wider(names_from = partner_type, values_from = cumulative) %>%
  group_by(fiscal_year) %>%
  mutate(Total = International + Local,
         share = Local / Total)  %>%
  pivot_longer(cols = International:Total, names_to = "partner_type") %>% 
  ungroup() %>% 
  filter(partner_type %in% c("Local"), fiscal_year == metadata_msd$curr_fy) %>% 
  pivot_wider(names_from = "partner_type", values_from = "value")
```

```{r gend_tbl, echo=FALSE, include = TRUE, warning=FALSE, message=FALSE, fig.align = 'left'}
qc_gend_lp %>% 
  select(-share) %>% 
  left_join(gbv_val_clean) %>% 
  rename(Total = cumulative) %>% 
  mutate(lp_share = Local/Total) %>% 
  gt() %>% 
  fmt_number(columns = c(5,6), 
             decimals = 0) %>% 
  fmt_percent(columns = c(7),
              decimals = 1) %>%
  grand_summary_rows(
    columns = c(5,6),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #33: GEND_GBV" %>% toupper())) %>% 
  gt_theme_nytimes() 

```
