---
title: "HFAC Briefing"
author: "Karishma Srikanth & Nada Petrovic"
date: "06/05/2023"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

  library(tidyverse)
  library(glitr)
  library(glamr)
  library(gophr)
  library(extrafont)
  library(scales)
  library(tidytext)
  library(patchwork)
  library(ggtext)
  library(glue)
  library(janitor)
  library(lubridate)
  library(googlesheets4)
```

# Modified Annual TPs to address HFAC request

```{r cars, include=FALSE, echo=FALSE}
#Current MSD
msd_path <- si_path() %>% 
  return_latest("MER_Structured_Datasets_OU_IM_FY21-23_20230512_v1_1")

get_metadata(msd_path)

df <- read_psd(msd_path)

# load source functions
#source("Scripts/archive/99_utilities.R")

#LOCAL PARTNER
#Read in the google sheet hyperfile with local partner
sheet_id <- "1MQviknJkJDttGdNEJeNaYPKmHCw6BqPuJ0C5cslV5IE"

df_partner <- read_sheet(sheet_id, sheet = "MechID-PartnerType", range = "A:B") %>% 
  clean_names() %>% 
  rename(mech_code = mechanism_id) %>% 
  mutate(mech_code = as.character(mech_code),
         partner_type = case_when(partner_type == "Regional" ~ "Local",
                                  partner_type == "TBD Local" ~ "Local", TRUE ~ partner_type))


#clean number function
clean_number <- function(x, digits = 0){
  dplyr::case_when(x >= 1e9 ~ glue("{round(x/1e9, digits)} billion"),
                   x >= 1e6 ~ glue("{round(x/1e6, digits)} million"),
                   x >= 1e3 ~ glue("{round(x/1e3, digits)} thousand"),
                   TRUE ~ glue("{x}"))
}

load_secrets()

```

## Topline Achievements

```{r tx_curr, echo=FALSE, include = FALSE}

 # Derived from agitprop/Scripts/26_treat_qtr_historic-txcurr-trends.R
  
  #create a PEPFAR duplicate and aggregate up to country/global level
  df_tx <- df %>%
    bind_rows(df %>% mutate(funding_agency = "PEPFAR")) %>% 
    filter(indicator %in% c("TX_CURR", "TX_NEW"),
           standardizeddisaggregate == "Total Numerator") %>% 
    group_by(fiscal_year, funding_agency, indicator) %>% 
    summarise(value = sum(cumulative, na.rm = TRUE)) %>% 
    ungroup() %>% 
    filter(funding_agency %in% c("PEPFAR", "USAID"),
           value != 0) %>% 
    arrange(indicator, funding_agency, fiscal_year) %>% 
    mutate(source = "MSD")

 df_tx <- df_tx %>% 
    filter(fiscal_year == metadata$curr_fy,
           indicator == "TX_CURR") %>%
    select(fiscal_year, indicator, funding_agency, value) %>% 
    pivot_wider(names_from = funding_agency) %>% 
    mutate(usaid_share = USAID/PEPFAR)  
  
  tx_val <<- df_tx %>% 
    mutate(USAID_Lab = USAID %>% clean_number()) %>%
    pull(USAID_Lab) 


```
```{r VLS, echo=FALSE, include = FALSE}

# derived from agitprop/Scripts/42_trend_ou_viral_load_coverage.R

df_vls_global <- df %>% 
    clean_indicator() %>% 
    clean_agency() %>% 
    clean_countries(colname = "country") %>% 
   filter(
      fiscal_year != metadata$curr_fy + 1,
      funding_agency == "USAID",
      indicator %in% c("TX_CURR", "TX_PVLS_D", "TX_PVLS"),
      standardizeddisaggregate %in% c("Total Numerator", "Total Denominator")
    ) %>% 
    group_by(fiscal_year, funding_agency, country, indicator) %>% 
    summarise(across(starts_with("qtr"), sum, na.rm = T), .groups = "drop") %>% 
    reshape_msd() %>% 
    select(-period_type) %>% 
    pivot_wider(names_from = indicator, values_from = value) %>% 
    group_by(funding_agency, country) %>% 
    mutate(TX_CURR_LAG2 = lag(TX_CURR, 2, order_by = period)) %>% 
    relocate(TX_CURR_LAG2, .before = TX_CURR) %>% 
    filter(!is.na(TX_PVLS_D), TX_PVLS_D > 0) %>% 
    group_by(period, funding_agency) %>% 
    summarise(across(starts_with("TX"), sum, na.rm = T), .groups = "drop") %>% 
    mutate(VLC = TX_PVLS_D / TX_CURR_LAG2,
           VLS = TX_PVLS/TX_PVLS_D)

  df_tx_ou <- df %>% 
    clean_indicator() %>% 
    clean_agency() %>% 
    clean_countries(colname = "country") %>% 
    filter(
      fiscal_year != metadata$curr_fy + 1,
      funding_agency == "USAID",
      indicator %in% c("TX_CURR", "TX_PVLS_D", "TX_PVLS"),
      standardizeddisaggregate %in% c("Total Numerator", "Total Denominator")
    ) %>% 
    group_by(fiscal_year, funding_agency, country, indicator) %>% 
    summarise(across(starts_with("qtr"), sum, na.rm = T), .groups = "drop") %>% 
    reshape_msd() %>% 
    dplyr::select(-period_type) %>% 
    pivot_wider(names_from = indicator, values_from = value) %>% 
    group_by(funding_agency, country) %>% 
    mutate(TX_CURR_LAG2 = lag(TX_CURR, 2, order_by = period),
           VLC = TX_PVLS_D / TX_CURR_LAG2,
           VLS = TX_PVLS/TX_PVLS_D) %>% 
    ungroup() %>% 
    relocate(TX_CURR_LAG2, .before = TX_CURR) %>% 
    filter(!is.na(TX_PVLS_D), TX_PVLS_D > 0)

glbl_vls <<- df_vls_global %>%
  filter(period == metadata$curr_pd) %>%
  mutate(VLS_pct = percent(VLS)) %>%
  pull(VLS_pct)

meet_90_vls <<- df_tx_ou %>%
  filter(period == metadata$curr_pd, VLS >= .9) %>%
  distinct(country) %>% nrow()

meet_95_vls <<- df_tx_ou %>%
  filter(period == metadata$curr_pd, VLS >= .95) %>%
  distinct(country) %>% nrow()

```


Results through **`r metadata$curr_pd`**

1. Through PEPFAR, the U.S. government has saved more than 21 million lives, prevented millions of HIV infections, and helped 20 countries with HIV burden achieve epidemic control of HIV or reach UNAIDS treatment targets. Country leadership on sustaining the HIV response will require bold political, programmatic, and financial leadership.

2. Currently, USAID provides over **`r tx_val`** clients with life-saving HIV-treatment. **`r glbl_vls`** of these beneficiaries who received a viral load test are virally suppressed, meaning that they can live longer, healthier lives and the virus can no longer be transmitted.
