---
title: "2024 WAD TPs - QC"
format: docx
editor: visual
---

# Purpose

Last Updated: `r lubridate::today()`

Previously used for: Data for World AIDS Day 2023, 2023 OHA Top Talking Points, PPR

Audience: The public, Agency stakeholders, partners, HIV/AIDS stakeholders/activists, the development community

Purpose: The purpose of this document is to create one place for the data points we will be sharing on World AIDS Day. On World AIDS Day we will be sharing (not a complete list): 

-   an Agency notice, 

-   an Agency-wide social media toolkit (and a USAID GH specific toolkit), 

-   a press statement from A/AID,

-   and we will be updating our OHA webpage on World AIDS Day [last year's version here](https://www.usaid.gov/world-aids-day).

```{r libraries}
#| echo: false
#| include: true
#| warning: false
#| message: false

#knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

  library(tidyverse)
  library(glitr)
  library(glamr)
  library(gophr)
  library(extrafont)
  library(scales)
  library(tidytext)
  library(patchwork)
  library(ggtext)
  library(glue)
  library(janitor)
  library(lubridate)
  library(googlesheets4)
  library(gt)
  library(gtExtras)

```

```{r load}
#| echo: false
#| include: true
#| warning: false
#| message: false

# SI specific paths/functions  
load_secrets()

# MSD
filepath <- si_path() %>% 
  return_latest("MER_Structured_Datasets_OU_IM_FY22")

# Grab metadata
metadata <- get_metadata(filepath) 

# MSD PREP --------------------------------------

#read MSD
df_msd <- read_psd(filepath)


#LOCAL PARTNER DATA PREP --------------------------------------------------------

#Read in the google sheet hyperfile with local partner
sheet_id <- "1tGk1TR8l3WacR8qMIK0AQvFynABijAaLHeIctE1nUoM"

df_partner <- read_sheet(sheet_id, sheet = "MechID-PartnerType") %>% 
  janitor::clean_names() %>% 
  select(mechanism_id, pepfar_local_partner_status) %>% 
  rename(mech_code = mechanism_id,
         partner_type = pepfar_local_partner_status) %>% 
  mutate(mech_code = as.character(mech_code))
         # ,
         # partner_type = case_when(partner_type == "Regional" ~ "Local",
         #                          partner_type == "TBD Local" ~ "Local", TRUE ~ partner_type))


# DREAMS DATA PREP --------------------------------------------------------------

# Google sheet ID for DREAMS DSNU list
dsnu_g_id <- "1oJZyXtvExN9UoQ6krFHERMoXEgCQgsHLf0tVIbmsi_o"

#MSD path
dsnu_msd_path <- si_path() %>% 
  return_latest("PSNU_IM_DREAMS_FY22")

#import DREAMS DSNU crosswalk
#recommend adding psnu_uids to this sheet
dsnu_list <- read_sheet(dsnu_g_id, sheet = "DSNU-Agency-All-FY") %>% 
  janitor::clean_names() %>% 
  select(ou, cop22_psnu, dreams_dsnu, cop_23_fy24_implementing_agency)

#rename names
names(dsnu_list)<-c("operatingunit", "psnu", "dsnu", "agencies_FY24")

#import msd and filter to current year
df_msd_dreams <- read_psd(dsnu_msd_path) %>% 
  filter(fiscal_year == metadata$curr_fy)

msd_dsnu_xwalk <- df_msd_dreams %>% 
  count(operatingunit, operatingunituid, psnu, psnuuid, cop22_psnu, cop22_psnuuid, dsnu, dsnuuid) %>% 
  mutate(raised_lvl = ifelse(psnuuid != cop22_psnuuid, TRUE, FALSE)) %>% 
  mutate(dsnu_new = case_when(operatingunit != "Uganda" & raised_lvl == TRUE ~ cop22_psnu,
                              TRUE ~ dsnu),
         dsnuuid_new = case_when(operatingunit != "Uganda" & raised_lvl == TRUE ~ cop22_psnuuid,
                                 TRUE ~ dsnuuid))

#join the new xwalk back to msd and then join the internal dsnu list to the msd
df_dreams_all <- df_msd_dreams %>% 
  left_join(msd_dsnu_xwalk) %>% 
  left_join(dsnu_list, by=c("operatingunit", "cop22_psnu" = "psnu", "dsnu_new"= "dsnu"))

# HRH -----------------------------------------------------------------------

df_hrh <- si_path() %>% 
  return_latest("HRH") %>% 
  read_psd()

```

```{r functions}
#| echo: false
#| include: true
#| warning: false
#| message: false

# FUNCTIONS --------------------------------------------------------------------

#clean number function
clean_number <- function(x, digits = 0){
  dplyr::case_when(x >= 1e9 ~ glue("{round(x/1e9, digits)} billion"),
                   x >= 1e6 ~ glue("{round(x/1e6, digits)} million"),
                   x >= 1e3 ~ glue("{round(x/1e3, digits)} thousand"),
                   TRUE ~ glue("{x}"))
}


#LP Function
#result_type = value or share
grab_lp_results_clean <- function(df, indic, result_type) {
  
  df_munge <- suppressMessages(df %>% 
                                 # bind_rows(df_arch) %>%
                                 left_join(df_partner, by = c("mech_code")) %>%
                                 filter(funding_agency == "USAID",
                                        indicator == indic,
                                        standardizeddisaggregate == "Total Numerator") %>% 
                                 group_by(funding_agency, fiscal_year, indicator, partner_type) %>% 
                                 summarise(across(cumulative, sum, na.rm = TRUE)) %>%
                                 ungroup() %>% 
                                 filter(partner_type != "TBD") %>%
                                 #filter(fiscal_year != 2022) %>%
                                 pivot_wider(names_from = partner_type, values_from = cumulative) %>%
                                 group_by(fiscal_year) %>%
                                 mutate(Total = International + Local,
                                        share = Local / Total)  %>%
                                 pivot_longer(cols = International:Total, names_to = "partner_type"))
  
  title_info_lp <- df_munge %>% 
    filter(partner_type == "Local", fiscal_year == metadata$curr_fy) %>% 
    select(fiscal_year, indicator, value, share) %>% 
    mutate(
      value = value %>% clean_number(1),#change to 1 if you want 1 decimal accuracy
      share = percent(round(share, 2))) %>% 
    pull(result_type)
  
  return(title_info_lp)
}

grab_lp_results <- function(df, indic, result_type) {
  
  df_munge <- suppressMessages(df %>% 
                                 # bind_rows(df_arch) %>%
                                 left_join(df_partner, by = c("mech_code")) %>%
                                 filter(funding_agency == "USAID",
                                        indicator == indic,
                                        standardizeddisaggregate == "Total Numerator") %>% 
                                 group_by(funding_agency, fiscal_year, indicator, partner_type) %>% 
                                 summarise(across(cumulative, sum, na.rm = TRUE)) %>%
                                 ungroup() %>% 
                                 filter(partner_type != "TBD") %>%
                                 #filter(fiscal_year != 2022) %>%
                                 pivot_wider(names_from = partner_type, values_from = cumulative) %>%
                                 group_by(fiscal_year) %>%
                                 mutate(Total = International + Local,
                                        share = Local / Total)  %>%
                                 pivot_longer(cols = International:Total, names_to = "partner_type"))
  
  title_info_lp <- df_munge %>% 
    filter(partner_type == "Local", fiscal_year == metadata$curr_fy) %>% 
    select(fiscal_year, indicator, value, share) %>% 
    mutate(
      value = value,#change to 1 if you want 1 decimal accuracy
      share = percent(round(share, 2))) %>% 
    pull(result_type) 
  
  return(title_info_lp)
}

#key pops function clean
pull_kp_clean <- function(df, indic) {
  kp_val <- df %>%
    filter(funding_agency == "USAID",
           str_detect(standardizeddisaggregate, "KeyPop(?!\\/Status)"),
           indicator %in% c("PrEP_NEW", "TX_CURR", "KP_PREV"),
           fiscal_year == metadata$curr_fy) %>%
    count(indicator, wt = cumulative) %>%
    mutate(n = clean_number(n)) %>% 
    pivot_wider(names_from = indicator, values_from = "n") %>% 
    pull(indic)
  
  return(kp_val)
}

#key pops function complete
pull_kp <- function(df, indic) {
  kp_val <- df %>%
    filter(funding_agency == "USAID",
           str_detect(standardizeddisaggregate, "KeyPop(?!\\/Status)"),
           indicator %in% c("PrEP_NEW", "TX_CURR", "KP_PREV"),
           fiscal_year == metadata$curr_fy) %>%
    count(indicator, wt = cumulative) %>%
    pivot_wider(names_from = indicator, values_from = "n") %>% 
    pull(indic) 
  
  return(kp_val)
}

```

### TP #1: FY24 PrEP_NEW

```{r prep_tp}
#| message: false
#| warning: false

prep_val_qc <- df_msd %>%
  filter(funding_agency == "USAID",
         standardizeddisaggregate %in% c("Total Numerator"),
         indicator %in% c("PrEP_NEW"),
         fiscal_year == metadata$curr_fy) %>%
  group_by(fiscal_year, funding_agency, indicator) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T)) %>% 
  ungroup()

```
```{r prep_tp_qc}
#| echo: false
#| include: true
#| warning: false
#| message: false

prep_val_qc %>% 
  #filter(`OU category` != "Total") %>% 
  gt::gt() %>% 
  fmt_number(columns = c(4), 
             decimals = 0) %>% 
  tab_header(
    title = glue("TP #1: USAID PrEP_NEW FY24" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

### TP #1: FY24 PrEP_NEW Countries

```{r prep_country_tp}
#| message: false
#| warning: false

prep_country_qc <- df_msd %>% 
  filter(funding_agency == "USAID",
         standardizeddisaggregate %in% c("Total Numerator"),
         indicator %in% c("PrEP_NEW"),
         fiscal_year == metadata$curr_fy) %>%
  distinct(fiscal_year, indicator, country) %>% 
  select(country) %>%
  mutate(id = row_number()) %>% 
  mutate(country_new = glue::glue("{id}. {country}")) %>% 
  select(country_new) %>%
  pull() 

```

```{r prep_country_tp_qc}
#| echo: false
#| include: true
#| warning: false
#| message: false

prep_country_qc
```

### TP #2: FY24 PrEP_NEW Local Partner Share

```{r prep_lp_tp}
#| message: false
#| warning: false
#| 
prep_lp_qc <- df_msd %>% 
  left_join(df_partner, by = c("mech_code")) %>%
  filter(funding_agency == "USAID",
         indicator == "PrEP_NEW",
         standardizeddisaggregate == "Total Numerator") %>%
  group_by(funding_agency, fiscal_year, indicator, partner_type) %>% 
  summarise(across(cumulative, sum, na.rm = TRUE)) %>%
  ungroup() %>% 
  filter(!str_detect(partner_type, "TBD")) %>%
  pivot_wider(names_from = partner_type, values_from = cumulative) %>%
  group_by(fiscal_year) %>%
  
  mutate(Total = International + Local)  %>%
  ungroup() %>% 
  pivot_longer(cols = International:Local, names_to = "partner_type") %>% 
  mutate(share = value / Total) %>% 
  filter(fiscal_year == metadata$curr_fy) %>% 
  select(fiscal_year, funding_agency, partner_type, indicator, value, share) 
  
```

```{r prep_lp_tp_qc}
#| echo: false
#| include: true
#| warning: false
#| message: false

prep_lp_qc %>% 
  #filter(`OU category` != "Total") %>% 
  gt::gt() %>% 
  fmt_number(columns = c(5), 
             decimals = 0) %>% 
  fmt_percent(columns = c(6),
              decimals = 0) %>% 
  tab_header(
    title = glue("TP #2: USAID Local Partner Share FY24" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

### TP #3: FY24 PrEP_NEW for KP and AGYW

```{r prep_kp_agyw_tp}
#| message: false
#| warning: false

kp_val_qc <- df_msd %>%
    filter(funding_agency == "USAID",
           str_detect(standardizeddisaggregate, "KeyPop(?!\\/Status)"),
           indicator %in% c("PrEP_NEW"),
           fiscal_year == metadata$curr_fy) %>%
  mutate(group= "KP") %>% 
    count(indicator,fiscal_year,funding_agency, group, wt = cumulative) %>%
    pivot_wider(names_from = indicator, values_from = "n") 


prep_agyw_val_qc <- df_msd %>%
  filter(funding_agency == "USAID",
         standardizeddisaggregate %in% c("Age/Sex"),
         indicator %in% c("PrEP_NEW"),
         fiscal_year == metadata$curr_fy,
         sex == "Female",
         age_2019 %in% c("10-14","15-19", "20-24")
          ) %>%
  mutate(group = "AGYW 10-24") %>% 
  group_by(indicator, fiscal_year, funding_agency, group) %>% 
  summarise(across(starts_with("cumulative"), sum, na.rm = T)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = indicator, values_from = "cumulative") 

prep_pp_qc <- bind_rows(kp_val_qc, prep_agyw_val_qc)

```

```{r prep_kp_agyw_qc}
#| echo: false
#| include: true
#| warning: false
#| message: false

prep_pp_qc %>% 
   gt::gt() %>% 
  fmt_number(columns = c(4), 
             decimals = 0) %>% 
  tab_header(
    title = glue("TP #3: USAID KP and AGYW PrEP_NEW FY24" %>% toupper())) %>% 
  gt_theme_nytimes() 

```


## DREAMS Talking Points

### TP #4: FY24 USAID AGYW_PREV for Women 10-29


```{r dreams_tp}
#| message: false
#| warning: false
#| 
# DREAMS ----------------------------------------------------------

#For total DREAMS programming, use all 4 disaggs + ages 10-29
viz_package <- df_dreams_all %>%
  filter(indicator=="AGYW_PREV",
         standardizeddisaggregate %in% c("Age/Sex/Time/Complete+",
                                         "Age/Sex/Time/Complete",
                                         "Age/Sex/Time/Started",
                                         "Age/Sex/Time/Incomplete"),
         numeratordenom=="D",
         age_2019 %in% c("10-14","15-19","20-24","25-29"),
         sex=="Female")

#service type disagg
viz_service<- df_dreams_all %>% filter(indicator=="AGYW_PREV", 
                                       standardizeddisaggregate %in% c("ComprehensiveEconomicStrengthening",
                                                                       "EducationSupport"),
                                       numeratordenom=="D")

#create shares by USAID and non-USAID contributing DSNUs
dreams_overall_val_qc <- viz_package %>% 
  mutate(usaid_dsnu = ifelse(str_detect(agencies_FY24, "USAID"), "USAID", "non-USAID")) %>% 
  group_by(indicator, fiscal_year, usaid_dsnu) %>% 
  summarize(across(cumulative, \(x) sum(x, na.rm = TRUE)),.groups="drop") %>% 
  pivot_wider(names_from = "usaid_dsnu", values_from = 'cumulative') %>% 
  mutate(total = USAID + `non-USAID`) %>% 
  pivot_longer(cols = c(3:4), names_to = "funding_agency") %>% 
  mutate(share = value / total) %>% 
  select(-c(total))

```

```{r dreams_qc}
#| echo: false
#| include: true
#| warning: false
#| message: false

dreams_overall_val_qc %>% 
  gt() %>% 
  fmt_number(columns = c(4), 
             decimals = 0) %>% 
  fmt_percent(columns = c(5),
              decimals = 0) %>% 
  grand_summary_rows(
    columns = where(is.numeric),
    fns = list(
      Overall = ~ sum(., na.rm = T)
    ),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_header(
    title = glue("TP #4: USAID AGYW_PREV for Women 10-29" %>% toupper())) %>% 
  gt_theme_nytimes() 

```

### TP #5: FY24 USAID DREAMS Mentors (HRH)

```{r dreams_hrh_tp}
#| message: false
#| warning: false

dreams_mentors_qc <- df_hrh %>% 
  filter(str_detect(employment_title, "DREAMS Mentor"),
         funding_agency == "USAID",
         fiscal_year == metadata$curr_fy) %>% 
  group_by(fiscal_year, funding_agency, employment_title) %>% 
  summarize(across(individual_count, \(x) sum(x, na.rm = TRUE)),.groups="drop") 

```

```{r dreams_hrh_qc}
#| echo: false
#| include: true
#| warning: false
#| message: false


dreams_mentors_qc %>% 
  gt() %>% 
  fmt_number(columns = c(4), 
             decimals = 0) %>% 
  # fmt_percent(columns = c(5),
  #             decimals = 0) %>% 
  # grand_summary_rows(
  #   columns = where(is.numeric),
  #   fns = list(
  #     Overall = ~ sum(., na.rm = T)
  #   ),
  #   formatter = fmt_number,
  #   decimals = 0
  # ) %>% 
  tab_header(
    title = glue("TP #5: FY24 USAID DREAMS Mentors" %>% toupper()),
    subtitle = "Source: FY24Q4i HRH Dataset") %>% 
  gt_theme_nytimes() 
```

## Local Partner Talking Points

### TP #5: FY24 USAID Local Partner Share for HTS and TX_CURR


```{r local_partner_tp}
#| message: false
#| warning: false

hts_tx_lp_qc <- df_msd %>% 
  left_join(df_partner, by = c("mech_code")) %>%
  filter(funding_agency == "USAID",
         indicator %in% c("HTS_TST", "TX_CURR"),
         standardizeddisaggregate == "Total Numerator") %>%
  group_by(funding_agency, fiscal_year, indicator, partner_type) %>% 
  summarise(across(cumulative, sum, na.rm = TRUE)) %>%
  ungroup() %>% 
  filter(!str_detect(partner_type, "TBD")) %>%
  pivot_wider(names_from = partner_type, values_from = cumulative) %>%
  group_by(fiscal_year) %>%
  mutate(Total = International + Local)  %>%
  ungroup() %>% 
  pivot_longer(cols = International:Local, names_to = "partner_type") %>% 
  mutate(share = value / Total) %>% 
  filter(fiscal_year == metadata$curr_fy) %>% 
  select(fiscal_year, funding_agency, indicator, partner_type, value, share)
```

```{r local_partner_qc}
#| echo: false
#| include: true
#| warning: false
#| message: false

hts_tx_lp_qc %>% 
  #filter(`OU category` != "Total") %>% 
  gt::gt() %>% 
  fmt_number(columns = c(5), 
             decimals = 0) %>% 
  fmt_percent(columns = c(6),
              decimals = 0) %>% 
  tab_header(
    title = glue("TP #6: FY24 USAID Local Partner Share for HTS and TX_CURR" %>% toupper())) %>% 
  gt_theme_nytimes() 
```

