---
title: "2023 World AIDS Day TPs"
author: "Karishma Srikanth"
date: "2023-11-28"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

  library(tidyverse)
  library(glitr)
  library(glamr)
  library(gophr)
  library(extrafont)
  library(scales)
  library(tidytext)
  library(patchwork)
  library(ggtext)
  library(glue)
  library(janitor)
  library(lubridate)
  library(googlesheets4)
```

# Purpose

Last Updated: 11/8/23

Previously used for: Data for World AIDS Day 2022, 2022 OHA Top Talking Points, PPR

Audience: The public, Agency stakeholders, partners, HIV/AIDS stakeholders/activists, the development community

Purpose: The purpose of this document is to create one place for the data points we will be sharing on World AIDS Day. On World AIDS Day we will be sharing (not a complete list): 

-   an Agency notice, 

-   an Agency-wide social media toolkit (and a USAID GH specific toolkit), 

-   a press statement from A/AID,

-   and we will be updating our OHA webpage on World AIDS Day [last year's version here](https://www.usaid.gov/world-aids-day).

These points will be repeated throughout these deliverables and language may vary slightly from what is included below. The USG World AIDS Day theme is "World AIDS Day 35 - Remember and Commit" (WAD 35 means 35 years of World AIDS Day). The UNAIDS theme is "[Let Communities Lead"](https://www.unaids.org/en/2023-world-aids-day#:~:text=Let%20Communities%20Lead%20%2D%20World%20AIDS,progress%20in%20the%20HIV%20response.)

Our plan is to share the Agency social media toolkit by Nov 23.

The clearance timeline is:

-   Division chiefs to clear by COB Tuesday Nov 14th
-   Front Office to clear by COB Friday November 17th
-   Nov 21st - SIEI & comms to address any outstanding issues/comments from clearance process
-   Nov 24nd - SIEI to plug in numbers for Q4
-   Nov 28rd - Social Media tool kit is released

```{r cars, include=FALSE, echo=FALSE}

  # SI specific paths/functions  
    load_secrets()

  # MSD
    filepath <- si_path() %>% 
      return_latest("MER_Structured_Datasets_OU_IM_FY21-24_20231114_v1_1")
    
      # Grab metadata
    get_metadata(filepath) 
  

  #read MSD
  df_msd <- read_psd(filepath)
    
  #read NGA genie file
  df_nga <- si_path() %>% 
    return_latest("All indicators_FY23Q4_Nigeria") %>% 
    read_csv()
  
  names_to_keep <- names(df_msd)
  setdiff(names(df_msd), names(df_nga)) #only qtrs
  setdiff(names(df_nga), names(df_msd))
  
  #select all names from NGA file that match MSD (minus qtr1-qtr4)
  df_nga_new <- df_nga %>% 
    select(any_of(names_to_keep))
  
#rbind together
  df <- df_msd %>% 
    filter(operatingunit != "Nigeria") %>% 
    select(-c(qtr1:qtr4)) %>% 
    rbind(df_nga_new)

# load source functions
#source("Scripts/archive/99_utilities.R")

#LOCAL PARTNER
#Read in the google sheet hyperfile with local partner
sheet_id <- "1MQviknJkJDttGdNEJeNaYPKmHCw6BqPuJ0C5cslV5IE"

df_partner <- read_sheet(sheet_id, sheet = "MechID-PartnerType", range = "A:B") %>% 
  clean_names() %>% 
  rename(mech_code = mechanism_id) %>% 
  mutate(mech_code = as.character(mech_code),
         partner_type = case_when(partner_type == "Regional" ~ "Local",
                                  partner_type == "TBD Local" ~ "Local", TRUE ~ partner_type))


#clean number function
clean_number <- function(x, digits = 0){
  dplyr::case_when(x >= 1e9 ~ glue("{round(x/1e9, digits)} billion"),
                   x >= 1e6 ~ glue("{round(x/1e6, digits)} million"),
                   x >= 1e3 ~ glue("{round(x/1e3, digits)} thousand"),
                   TRUE ~ glue("{x}"))
}

load_secrets()

```

## Topline Achievements

```{r tx_curr, echo=FALSE, include = FALSE}

 # Derived from agitprop/Scripts/26_treat_qtr_historic-txcurr-trends.R
  
  #create a PEPFAR duplicate and aggregate up to country/global level
  df_tx <- df %>%
    bind_rows(df %>% mutate(funding_agency = "PEPFAR")) %>% 
    filter(indicator %in% c("TX_CURR", "TX_NEW"),
           standardizeddisaggregate == "Total Numerator") %>% 
    group_by(fiscal_year, funding_agency, indicator) %>% 
    summarise(value = sum(cumulative, na.rm = TRUE)) %>% 
    ungroup() %>% 
    filter(funding_agency %in% c("PEPFAR", "USAID"),
           value != 0) %>% 
    arrange(indicator, funding_agency, fiscal_year) %>% 
    mutate(source = "MSD")

 df_tx <- df_tx %>% 
    filter(fiscal_year == metadata$curr_fy,
           indicator == "TX_CURR") %>%
    select(fiscal_year, indicator, funding_agency, value) %>% 
    pivot_wider(names_from = funding_agency) %>% 
    mutate(usaid_share = USAID/PEPFAR)  
  
  tx_val <<- df_tx %>% 
    mutate(USAID_Lab = USAID %>% clean_number()) %>%
    pull(USAID_Lab) 


```

```{r VLS, echo=FALSE, include = FALSE}

# derived from agitprop/Scripts/42_trend_ou_viral_load_coverage.R

df_vls_global <- df %>% 
    clean_indicator() %>% 
    clean_agency() %>% 
    clean_countries(colname = "country") %>% 
    filter(
      fiscal_year != metadata$curr_fy + 1,
      funding_agency == "USAID",
      indicator %in% c("TX_PVLS_D", "TX_PVLS"),
      standardizeddisaggregate %in% c("Total Numerator", "Total Denominator")
    ) %>% 
    group_by(fiscal_year, funding_agency, country, indicator) %>% 
    summarise(across(cumulative, sum, na.rm = T), .groups = "drop") %>% 
    # reshape_msd() %>% 
    # select(-period_type) %>% 
    pivot_wider(names_from = indicator, values_from = cumulative) %>% 
    group_by(funding_agency, country) %>% 
    # mutate(TX_CURR_LAG2 = lag(TX_CURR, 2, order_by = period)) %>% 
    # relocate(TX_CURR_LAG2, .before = TX_CURR) %>% 
    filter(!is.na(TX_PVLS_D), TX_PVLS_D > 0) %>% 
    group_by(fiscal_year, funding_agency) %>% 
    summarise(across(starts_with("TX"), sum, na.rm = T), .groups = "drop") %>% 
    mutate(VLS = TX_PVLS/TX_PVLS_D)

  df_tx_ou <- df %>% 
    clean_indicator() %>% 
    clean_agency() %>% 
    clean_countries(colname = "country") %>% 
    filter(
      fiscal_year != metadata$curr_fy + 1,
      funding_agency == "USAID",
      indicator %in% c("TX_PVLS_D", "TX_PVLS"),
      standardizeddisaggregate %in% c("Total Numerator", "Total Denominator")
    ) %>% 
    group_by(fiscal_year, funding_agency, country, indicator) %>% 
    summarise(across(cumulative, sum, na.rm = T), .groups = "drop") %>% 
    # reshape_msd() %>% 
    # dplyr::select(-period_type) %>% 
    pivot_wider(names_from = indicator, values_from = cumulative) %>% 
    group_by(funding_agency, country) %>% 
    mutate(VLS = TX_PVLS/TX_PVLS_D) %>% 
    ungroup() %>%
    filter(!is.na(TX_PVLS_D), TX_PVLS_D > 0)

  glbl_vls <<- df_vls_global %>%
    filter(fiscal_year == metadata$curr_fy) %>%
    mutate(VLS_pct = percent(VLS)) %>%
    pull(VLS_pct)
  
  meet_90_vls <<- df_tx_ou %>%
    filter(fiscal_year == metadata$curr_fy, VLS >= .9) %>%
    distinct(country) %>% nrow()
  
  meet_95_vls <<- df_tx_ou %>%
    filter(fiscal_year == metadata$curr_fy, VLS >= .95) %>%
    distinct(country) %>% nrow()

```

```{r HTS, echo=FALSE, include = FALSE}

  # derived from agitprop/Scripts/19_test_qtr-hts-trends.R

#HTS_TST dataset
munge_hts <- function(indic) {
  df_hts <- df %>% 
    filter(indicator == indic,
           standardizeddisaggregate == "Total Numerator") 
  
  df_hts <- suppressMessages(
    df_hts %>%
    bind_rows(df_hts %>% mutate(funding_agency = "PEPFAR")) %>%
    group_by(fiscal_year, funding_agency) %>%
    summarise(across(c(cumulative), sum, na.rm = TRUE)) %>%
    ungroup() %>%
    filter(funding_agency %in% c("PEPFAR", "USAID")) %>%
    mutate(source = "MSD") %>%
    rename(value = cumulative) %>%
    pivot_wider(names_from = funding_agency, values_from = value) %>%
    group_by(fiscal_year) %>%
    mutate(share = USAID / PEPFAR)  %>%
    pivot_longer(cols = PEPFAR:USAID, names_to = "funding_agency"))
  
  hts_val <- df_hts %>% 
    filter(fiscal_year == metadata$curr_fy,
           funding_agency == "USAID") %>% 
    select(fiscal_year, value, share) %>% 
    mutate(share = percent(round(share, 2)),
           value = round(value, 2),
           value_clean = value %>%  clean_number()) %>%
    pull(value_clean)
  
  
  return(hts_val)
}

hts_val <- munge_hts("HTS_TST")
hts_pos_val <- munge_hts("HTS_TST_POS")
```

```{r site counts, echo=FALSE, include = FALSE}

  # still figuring out API call for this TP - just run this script on your own

 # source("Scripts/xx_hss_ann_usaid-site-count.R")

```

```{r LP, echo=FALSE, include = FALSE}

#catch-22/Scripts/20211102_Local Partner Meeting/20211022_LP_TX_trends.R

#result_type = value or share
grab_lp_results <- function(indic, result_type) {
  
  df_munge <- suppressMessages(df %>% 
    # bind_rows(df_arch) %>%
    left_join(df_partner, by = c("mech_code")) %>%
    filter(funding_agency == "USAID",
           indicator == indic,
           standardizeddisaggregate == "Total Numerator") %>% 
    group_by(funding_agency, fiscal_year, indicator, partner_type) %>% 
    summarise(across(cumulative, sum, na.rm = TRUE)) %>%
    ungroup() %>% 
    filter(partner_type != "TBD") %>%
    #filter(fiscal_year != 2022) %>%
    pivot_wider(names_from = partner_type, values_from = cumulative) %>%
    group_by(fiscal_year) %>%
    mutate(Total = International + Local,
           share = Local / Total)  %>%
    pivot_longer(cols = International:Total, names_to = "partner_type"))
  
  title_info_lp <- df_munge %>% 
    filter(partner_type == "Local", fiscal_year == metadata$curr_fy) %>% 
    select(fiscal_year, indicator, value, share) %>% 
    mutate(
      value = value %>%  clean_number(0), #change to 1 if you want 1 decimal accuracy
      share = percent(round(share, 2))) %>% 
    pull(result_type)
  
  return(title_info_lp)
}

tx_lp_val <- grab_lp_results("TX_CURR", "value")
tx_lp_share <- grab_lp_results("TX_CURR", "share")
hts_lp_val <- grab_lp_results("HTS_TST", "value")


```

Results through **`r metadata$curr_pd`**

1.  *Currently, USAID provides over **`r tx_val`** people living with HIV with life-saving HIV-treatment. **`r glbl_vls`** of these beneficiaries who received a viral load test are virally suppressed, meaning that they can live longer, healthier lives and the virus can no longer be transmitted.*

2.  Over the last year, USAID has provided HIV testing services to over **`r hts_val`** people.

3.  Over the last year, USAID helped over **`r hts_pos_val`** people learn of their HIV positive status.

4.  Over the last year, USAID and its partners supported **`r meet_90_vls`** countries to reach over 90 percent viral load suppression among people living with HIV, and of those, **`r meet_95_vls`** countries reached over 95 percent viral load suppression.

5.  *USAID investments strengthen host country systems that drive responsive, resilient, and enduring health care through more than **10,000** HIV testing sites, over **9,000** HIV treatment sites, and over **3,000** sites that provide lab-based or point-of-care testing making critical services accessible in over 50 supported PEPFAR countries.*

6.  *Local partners are on the forefront of the HIV response, supporting **over half of all USAID-supported people living with HIV** with clinical services.*

7.  *In 2022, local partners helped nearly **`r hts_lp_val`** people learn their HIV status and provided over **`r tx_lp_val`** people with life-saving HIV medication, representing **`r tx_lp_share`** of all USAID-supported people living with HIV.*

8.  *Since 2015, USAID support has enabled nearly **three million infants** to be born HIV-free to mothers living with HIV.*

9.  *Prevention continues to be one of USAID's fastest growing programs, with nearly **3.3 million pre-exposure prophylaxis (PrEP) initiations** across 42 countries since FY16, over **8.8 voluntary male medical circumcisions (VMMC)** across 13 countries, and over **280M condoms** and **30M personal lubricants** procured for 30 countries in FY23.*

## **Achievements related to those who are most disproportionately impacted by HIV**

```{r kp prep, echo=FALSE, include = FALSE}


pull_kp <- function(indic) {
kp_val <- df %>%
    filter(funding_agency == "USAID",
           str_detect(standardizeddisaggregate, "KeyPop(?!\\/Status)"),
           indicator %in% c("PrEP_NEW", "TX_CURR", "KP_PREV"),
           fiscal_year == metadata$curr_fy) %>%
    count(indicator, wt = cumulative) %>%
  mutate(n = clean_number(n, 1)) %>% 
    pivot_wider(names_from = indicator, values_from = "n") %>% 
  pull(indic)

return(kp_val)
}

kp_prev_val <- pull_kp("KP_PREV")
kp_tx_val <- pull_kp("TX_CURR")
kp_prep_val <- pull_kp("PrEP_NEW")

  kp_tx_ou <- df %>%
    filter(funding_agency == "USAID",
           str_detect(standardizeddisaggregate, "KeyPop(?!\\/Status)"),
           indicator %in% c("TX_CURR"),
           fiscal_year == metadata$curr_fy) %>% 
    group_by(fiscal_year, funding_agency, country, indicator) %>% 
    summarise(across(cumulative, sum, na.rm = T), .groups = "drop") %>% 
   # reshape_msd() %>% 
    #filter(period == metadata$curr_pd) %>% 
    distinct(country) %>% 
    nrow() 


#AGYW prep
prep_agyw_val <- df %>%
  filter(funding_agency == "USAID",
         standardizeddisaggregate %in% c("Age/Sex"),
         indicator %in% c("PrEP_NEW"),
         fiscal_year == metadata$curr_fy,
         sex == "Female",
         age_2019 %in% c("15-19", "20-24")
         ) %>%
  count(indicator, wt = cumulative) %>%
  mutate(n = clean_number(n))

```

```{r ovc, echo=FALSE, include = FALSE}
ovc_serv_val <- df %>% 
    filter(indicator %in% c("OVC_SERV"),
           standardizeddisaggregate %in% c("Total Numerator"),
         #  trendscoarse == "<18",
           fiscal_year == metadata$curr_fy,
           funding_agency == "USAID") %>% 
    group_by(fiscal_year, funding_agency) %>% 
    summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
    ungroup() %>% 
    mutate(cumulative = clean_number(cumulative, 2)) %>% 
  pull(cumulative)
```

```{r gbv, echo=FALSE, include = FALSE}

# agitprop/Scripts/archive/13_gend_gbv_BAN_map.R

# for now, assume this is GEND_GBV total numerator - check with Allison S. tomorrow

gbv_val <- df %>% 
  filter(indicator == "GEND_GBV",
         standardizeddisaggregate == "Total Numerator",
         funding_agency == "USAID",
         fiscal_year == metadata$curr_fy) %>% 
    group_by(fiscal_year, funding_agency) %>% 
    summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
        ungroup() %>% 
    mutate(cumulative = clean_number(cumulative)) %>% 
  pull(cumulative)

#local partner - GBV
gbv_lp_pct <- grab_lp_results("GEND_GBV", "share")



```

```{r vmmc, echo=FALSE, include = FALSE}
vmmc_val <- df %>% 
  filter(indicator == "VMMC_CIRC",
         standardizeddisaggregate == "Total Numerator",
         funding_agency == "USAID",
         fiscal_year == metadata$curr_fy) %>% 
    group_by(fiscal_year, funding_agency) %>% 
    summarise(across(starts_with("cumulative"), sum, na.rm = T), .groups = "drop") %>% 
        ungroup() %>% 
    mutate(cumulative = clean_number(cumulative)) %>% 
  pull(cumulative)
```

8.  USAID initiated almost **`r kp_prep_val`** members of key populations (KP) and **`r prep_agyw_val`** adolescent girls and young women (AGYW) on Pre-Exposure Prophylaxis (PrEP) during the last year, expanding effective HIV prevention to key and priority populations. Together KP and AGYW constitute *more than half* of all USAID PrEP initiations, and KP and AGYW PrEP initiations grew quarter over quarter throughout the entire year.

9.  This year alone, more than **`r kp_tx_val`** members of key populations in **`r kp_tx_ou`** countries are receiving life-saving HIV medication through USAID support.

10. This year USAID and our partners supported almost **`r kp_prev_val`** members of key populations, including sex workers, transgender people, men who have sex with men and people who inject drugs, with HIV prevention services and supportive interventions, which improve their lives and the lives of their children.

11. In the last year, USAID supported nearly **`r ovc_serv_val`** orphans and vulnerable children (OVC) and their families affected by HIV. Services provided to OVC families included, among others, household economic strengthening, education support, parenting skills building, facilitating access to maternal and child healthcare, and HIV specific testing, care, and treatment.

